---
title: "punisheR"
author: "Jill Cates, Tariq Hassan, Avinash Prabhakaran"
date: "`r Sys.Date()`"
output: 
    github_document : default
    rmarkdown::html_vignette : default
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r setup, include = FALSE}

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

[punisheR](https://github.com/UBC-MDS/punisheR) is a package for feature and model selection in R. Specifically, this package implements tools for forward and backward model selection. In order to measure model quality during the selection procedures, we have also implemented the Akaike Information Criterion (AIC) and Bayesian Information Criterion (BIC).

## Functions included

The package contains two stepwise feature selection techniques:

- `forward()`: [Forward selection](https://en.wikipedia.org/wiki/Stepwise_regression#Main_approaches) starts with one feature and iteratively adds the features with the best scores using a model fit criterion. The process of adding features is repeated until either the maximum number of features (`n_features`) is reached or the change in score is less than the `min_change` threshold.


- `backward()`: [Backward selection/elimination](https://en.wikipedia.org/wiki/Stepwise_regression#Main_approaches) starts with all features and iteratively deletes features with the worst scores using model fit criterion. The process of deleting features is repeated until either the maximum number of features (`n_features`) is reached or the change in score is less than the `min_change` threshold.

Sources: https://en.wikipedia.org/wiki/Stepwise_regression

The package contains three metrics that evaluate model performance: 

- `aic()`: The [Akaike information criterion](https://en.wikipedia.org/wiki/Akaike_information_criterion) (AIC) adds a penalty term which penalizes more complex models. Its formal definition is:
$$-2\ln(L)+2*k $$
where $k$ is the number of features and $L$ is the maximized value of the likelihood function.


- `bic()`: The [Bayesian information criterion](https://en.wikipedia.org/wiki/Bayesian_information_criterion) adds a penality term which penalizes complex models to a greater extent than AIC. Its formal definition is:
$$-2*\ln(L)+\ln(n)*k$$
where $k$ is the number of features, $n$ is the number of observations, and $L$ is the maximized value of the likelihood function.

- `r_squared()`: The [coefficient of determination](https://en.wikipedia.org/wiki/Coefficient_of_determination) is the proportion of the variance in the response variable that can be predicted from the explanatory variable.

These three criteria measure the relative quality of models within `forward()` and `backward()` and can be configured using the `criterion` parameter. In general, having more parameters in your model increases prediction accuracy but is highly susceptible to overfitting. AIC and BIC add a penalty for the number of features in a model. The lower the AIC and BIC score, the better the model.

## How does punisheR fit into the existing R ecosystem?

In the R ecosystem, forward and backward selection are implemented in both the [olsrr](https://cran.r-project.org/web/packages/olsrr/)
and [MASS](https://cran.r-project.org/web/packages/MASS/MASS.pdf) packages. The former provides
[`ols_step_forward()`](https://www.rdocumentation.org/packages/olsrr/versions/0.4.0/topics/ols_step_forward) and
[`ols_step_backward()`](https://www.rdocumentation.org/packages/olsrr/versions/0.4.0/topics/ols_step_backward) for forward and backward stepwise selection, respectively. Both of these use p-value as a metric for feature selection. The latter, MASS, contains [`StepAIC()`](https://stat.ethz.ch/R-manual/R-devel/library/MASS/html/stepAIC.html), which is complete with three modes: forward, backward or both. Other packages that provide subset selection for regression models are [leaps](https://cran.r-project.org/web/packages/leaps/leaps.pdf) and [bestglm](https://cran.r-project.org/web/packages/bestglm/bestglm.pdf).



```{r}
library(knitr)
library(punisheR)
```


```{r}
#Loading the demo mtcars data
data <- mtcars_data()
X_train <- data[[1]]
y_train <- data[[2]]
X_val <- data[[3]]
y_val <- data[[4]]
```


## Forward Selection by specifying the number of features

###### Usage example with `aic` as criterion

```{r}
forward(X_train, y_train, X_val, y_val, min_change=NULL,
                    n_features=2, criterion='aic', verbose=FALSE)
```
When implementing forward selection on the mtcars dataset with `hp` as the explanatory variable , it returns a list of features that form the best model. In the above example, the desired number of features has been specified as 2 and the criterion being used is `aic`. The function returns a list of 2 features.

###### Usage example with `bic` as criterion

```{r}
forward(X_train, y_train, X_val, y_val, min_change=NULL,
                    n_features=3, criterion='bic', verbose=FALSE)
```

In the above example, the desired number of features has been specified as 3 and the criterion being used is `bic`. The function returns a list of 3 features.

###### Usage example with `r-squared` as criterion

```{r}
forward(X_train, y_train, X_val, y_val, min_change=NULL,
                    n_features=4, criterion='r-squared', verbose=FALSE)
```

In the above example, the desired number of features has been specified as 4 and the criterion being used is `r-squared`. The function returns a list of 4 features.


#### Forward Selection by specifying the smallest change in criterion 

```{r}
forward(X_train, y_train, X_val, y_val, min_change=0.5,
                    n_features=NULL, criterion='r-squared', verbose=FALSE)
```

In the example above, `forward` selction returns a list of 6 features when a minimum change of 0.5 is required in `r-squared` score for an additional feature to be selected.

**Note**: When using the criterion as `aic` or `bic`, the value for `min_change` should be carefully selected as `aic` and `bic` tends to have much larger values.



#### Backward Selection by specifying the number of features

```{r}
backward(X_train, y_train, X_val, y_val,
                     n_features=7, min_change=NULL, criterion='aic',
                     verbose=FALSE)
```

```{r}
backward(X_train, y_train, X_val, y_val,
                     n_features=7, min_change=NULL, criterion='bic',
                     verbose=FALSE)
```

```{r}
backward(X_train, y_train, X_val, y_val,
                     n_features=7, min_change=NULL, criterion='r-squared',
                     verbose=FALSE)
```

Similarly, for backward selection, the number of features are specified as 7 and the examples using all the three criterion are provided above.

#### Backward Selection by specifying the smallest change in criterion

```{r}
backward(X_train, y_train, X_val, y_val,
                     n_features=NULL, min_change=0.5, criterion='r-squared',
                     verbose=FALSE)
```

In the example above, `backward` selection returns a list of 10 features when the minimum change in the `r-squared` criterion is specified as 0.5.

#### AIC, BIC & $R^2$

```{r}
model <- lm(y_train ~ mpg + cyl + disp, data = X_train)

aic(model)
```

```{r}
bic(model)
```

When scoring the two the model using AIC and BIC, we can see that the penalty when using `bic` is greater than the penalty obtained using `aic`.

```{r}
r_squared(model, X_val, y_val)
```

The value returned by the function `r_squared()` will be between 0 and 1.
