---
title: "punisheR"
author: "Jill Cates, Tariq Hassan, Avinash Prabhakaran"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}

knitr::opts_chunk$set(eval = FALSE)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

PunisheR is a package for feature and model selection in R. Specifically, this package will implement tools for forward and backward model selection. In order to measure model quality during the selection procedures, we have also implemented the Akaike Information Criterion(AIC) and Bayesian Information Criterion(BIC).

## Functions included

The package contains two stepwise feature selection techniques:

- `forward()`: [Forward selection](https://en.wikipedia.org/wiki/Stepwise_regression#Main_approaches), which involves starting with no variables in the model, testing the addition of each variable using a chosen model fit criterion, adding the variable (if any) whose inclusion gives the most statistically significant improvement of the fit, and repeating this process until none improves the model to a statistically significant extent.

- `backward()`: [Backward selection/elimination](https://en.wikipedia.org/wiki/Stepwise_regression#Main_approaches), which involves starting with all candidate variables, testing the deletion of each variable using a chosen model fit criterion, deleting the variable (if any) whose loss gives the most statistically insignificant deterioration of the model fit, and repeating this process until no further variables can be deleted without a statistically significant loss of fit.

The package contains two metrics that evaluate model performance: 

- `aic()`: The [Akaike information criterion](https://en.wikipedia.org/wiki/Akaike_information_criterion) (AIC) is an estimator of the relative quality of statistical models for a given set of data.

- `bic()`: The [Bayesian information criterion](https://en.wikipedia.org/wiki/Bayesian_information_criterion) (BIC) or Schwarz criterion is a criterion for model selection among a finite set of models. It is based, in part, on the [likelihood function](https://en.wikipedia.org/wiki/Likelihood_function) and it is closely related to the [Akaike information criterion](https://en.wikipedia.org/wiki/Akaike_information_criterion) (AIC).


These two criteria will be used to measure the relative quality of models within `forward_selection()` and `backward_selection()`. In general, having more parameters in your model increases prediction accuracy but is highly susceptible to overfitting. AIC and BIC add a penalty for the number of features in a model. The penalty term is larger in BIC than in AIC. The lower the AIC and BIC score, the better the model.

## How does punisheR fit into the existing R ecosystem?

In the R ecosystem, forward and backward selection are implemented in both the [olsrr](https://cran.r-project.org/web/packages/olsrr/)
and [MASS](https://cran.r-project.org/web/packages/MASS/MASS.pdf) packages. The former provides
[`ols_step_forward()`](https://www.rdocumentation.org/packages/olsrr/versions/0.4.0/topics/ols_step_forward) and
[`ols_step_backward()`](https://www.rdocumentation.org/packages/olsrr/versions/0.4.0/topics/ols_step_backward) for
forward and backward stepwise selection, respectively. Both of these are p-value-based methods of feature selection. The latter, MASS, contains [`StepAIC()`](https://stat.ethz.ch/R-manual/R-devel/library/MASS/html/stepAIC.html),
which is complete with three modes: forward, backward or both. The selection procedure it uses is based on an
information criterion (AIC), as we intend ours to be.

```{r}
X_train <- data.frame(x1=rnorm(10), x2=rep(0,10), x3=rnorm(10), x4=rep(0, 10))
y_train <- rnorm(10)
X_val <- data.frame(x1=rnorm(10), x2=rep(0,10), x3=rnorm(10), x4=rep(0, 10))
y_val <- rnorm(10)
```



```{r}
forward(lm, X_train, y_train, X_val, y_val,
                     n_features=2, min_change=0.5, criterion='aic',
                     verbose=TRUE)
```

```{r}
forward(lm, X_train, y_train, X_val, y_val,
                     n_features=2, min_change=0.5, criterion='bic',
                     verbose=TRUE)
```

```{r}
backward(lm, X_train, y_train, X_val, y_val,
                     n_features=2, min_change=0.5, criterion='aic',
                     verbose=TRUE)
```

```{r}
backward(lm, X_train, y_train, X_val, y_val,
                     n_features=2, min_change=0.5, criterion='bic',
                     verbose=TRUE)
```

```{r}
aic(lm(y_train~X_train))
```


```{r}
bic(lm(y_train~X_train))
```


